% Title page
\DeclareTranslation{English}{thesis-institute}{Institute of Computer Science II -- Visual Computing}
\DeclareTranslation{English}{thesis-university}{University of Bonn}
\DeclareTranslation{English}{thesis-first-reviewer}{First Reviewer}
\DeclareTranslation{English}{thesis-second-reviewer}{Second Reviewer}

% Summary chapter
\DeclareTranslation{English}{thesis-summary}{Summary}

% Acknowledgments chapter
\DeclareTranslation{English}{thesis-acknowledgments}{Acknowledgments}

% Disclosure chapter
\DeclareTranslation{English}{thesis-disclosure}{Disclosure of Generative AI Usage}
\DeclareTranslation{English}{thesis-disclosure-statement}{%
In appreciation of the increasing applicability of Generative AI in academic work, I openly acknowledge its utilization in preparation of this thesis.
%
Generative AI tools were used to help reword and polish paragraphs to improve readability and ease of understanding.
%
Generative AI tools also helped in researching applicable libraries and technology, as well as interpreting various concepts covered here.
%
All work created or edited through Generative AI has been independently checked and approved by me as true and valid.
%
I claim sole responsibility for the content and attest that the end-product represents my own comprehension and effort.
%
No content or output from any Generative AI was used verbatim without critical evaluation and rewording.
%
All sources, human as well as AI-generated, have been cited where applicable and verified and checked on truthfulness.
%
Generative AI was used as a supportive tool only and never as a substitute for original research, analysis, or critical thinking involved in this thesis.
}
